---
title: "Replication of Study by Shavit et al. (2022, Work, Aging, and Retirement)"
author: "Enna Chen (ennachen@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---

<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

## Introduction

### The original study

This study examined how time horizon (unspecified or expanded) influence the motivation (demonstrated by choice from three work-related activities) of older and younger workers (age measured as a continuous variable).

Based on socioemotional selectivity theory (SST), they hypothesized that under unspecified time horizons, older workers prefer to engage in emotionally meaningful work activities more so than younger workers. They further hypothesized that when time horizons at work are expanded or limited, age differences are eliminated. Consistent with SST postulates, they found that age was associated with preferences for helping colleagues in the unspecified horizons condition, and that age differences were eliminated when time horizons were extended or limited. As hypothesized, expanding time horizons reduced employees’ likelihood of choosing to help colleagues. Contrary to their hypothesis, limiting time horizons also reduced the likelihood of choosing to help colleagues. 

### Justification

This paper is the latest publication (July 2022) from my lab that I am really intrigued by. It is along my line of research -- age, motivation, and prosociality. Additionally, it has a similar type of design as my intended first-year project (also potentially a life-span sample divided into three experimental conditions). I have never conducted a study with this type of design, so I hope to use this class project to practice for my own data collection and analysis. 

### Stimuli, procedures, challenges

Participants will be randomly assigned to one of two experimental conditions where time horizons at work were unspecified or expanded. Each condition includes a brief manipulation prompt, which had been validated in the original study. After reading the prompt, we ask participants in each condition to choose one of three work-related activities, each of which represent pursuit of a different type of goal. In the end, participants will provide information on their age, gender, and subjective socioeconomical status.

This Qualtrics survey would take approximately 2 minutes. The original paper has a sample of 555 adults aged 20 to 75 (M = 44.65, SD = 13.2). I will recruit a sample of 170 participants that matches with this age distribution by creating age buckets on Prolific. I will recruit only people who have a full-time. 

### Links

[Link to my repo](https://github.com/psych251/shavit2022)
[Link to the original paper](https://github.com/psych251/shavit2022/blob/main/original_paper/original_paper.pdf)
[Link to my pre-registration](https://osf.io/86p4e/)

## Methods

### Power Analysis 

I used z-tests: logistic regression - a priori power analysis. 
Original effect size: -.42
Odds ratio: 0.657
Pr H0: 0.33
alpha probability: 0.05
Effect sizes to achieve 80%, 90%, 95% power: 170, 232, 291
Considering our class funding, I decided to go with the 170 sample size that will achieve 80% power.

### Planned Sample

We will recruit 170 participants relatively evenly distributed from ages 20-75. In the paper, it was 20 to 75 years with a mean of 44.65, and SD of 13.2, and it was similar across conditions. They will all be full-time employees.

### Materials + Procedure

Participants were randomly assigned to one of two experimental conditions where time horizons at work were either unspecified or expanded. In the unspecified condition, participants were asked to imagine that they arrived at work one day and found that they had unscheduled time. Participants in the expanded time horizons condition were told that “the company would like you to stay with them considerably longer than you had planned.”

After reading the prompt, participants in each condition were asked to choose one of three work-related activities that they would pursue: “Help a colleague or a friend with a company project,” “Work on a project that may help you advance you career,” or “Work on a project that may lead the company in a new direction.” Helping colleagues represented pursuit of emotionally meaningful goals, whereas career-advancing and taking the company in new directions represented pursuit of exploratory goals (future payoffs and learning new information, respectively).

Finally, participants will provide background information - age, gender, subjective socioeconomic status, and employment status.

I have obtained the original questionnaire materials from the author.
[Link to my questionnaire](https://stanforduniversity.qualtrics.com/jfe/form/SV_0Utk18jvCwCk61w)

### Analysis Plan

**Clarify key analysis of interest here. You can also pre-specify additional analyses you plan to do.**  

The original paper did not mention data cleaning or exclusion rules. I would double check participants' answers to the age and employment status questions in Qualtrics to make sure to exclude those who are not 20-75 or full-time employees, in case the exclusion requirement from Prolific was not accurate.

The original paper included many covariates (e.g., Big Five Personality, Future Time Perspective, etc.). To be mindful of our class funding, I will only include basic demographic information (age, gender, employment status, and subjective socioeconomic status) as covariates.

Below are my three proposed statistical tests:

Key test:
1. The main test focuses on the difference (across ages) in the likelihood of choosing to help colleagues between the unspecified and the expanded time horizon conditions. I will use logistic regression to look at how the likelihood of deciding to help others (vs. exploratory goals) is associated with z-scored age, condition (unspecified vs. expanded time horizons), and their interaction. I will use two logistic regression models, one to set the unspecified time horizon condition as the reference group, and the other one to set the expanded time horizon condition as the reference group. Whether the replication is successful will be determined by whether there is an interaction effect of condition and age.

Exploratory tests:
1. Hypothesis 1 states that when time horizons are unspecified, older age is associated with a preference to help others over future-oriented options. Our first exploratory test focuses on the simple effect of age on the likelihood of choosing to help a colleague or a friend over any of the future-oriented options in the unspecified time horizon condition.
2. Hypothesis 2 states that when time horizons are expanded, age differences in work preferences are eliminated such that across all ages, future oriented options are preferred over helping others. Our second exploratory test focuses on the simple effect of age on the likelihood of choosing to help a colleague or a friend over any of the future-oriented options in the expanded time horizon condition.

### Differences from Original Study

The focal tests of the original paper for the three hypotheses are the simple effects of age on the likelihood of choosing to help a colleague or a friend over any of the future-oriented options, in each of the three conditions. Note that in this replication study, we only included two of the three conditions that had the largest effect. 

The original paper estimated the effects of age and time horizons condition using logistic regressions in which the intercept indicates a difference from a chance level of 33%. Although they did not expect to find significant Age x Condition interactions because they were not found in past research, they included them in the model to provide context for understanding the focal tests of age associations with preferences in each of the conditions. 

As a first step, they entered only age, condition, and their interaction term into the model. In a second step, to ensure any observed age effects are not caused by third variables, they included background variables found to be significantly associated with age in the current sample as control variables. To facilitate interpretation, they supplemented this approach with chi-square analysis. Because the results did not change after including the control variables, and that none of the covariates are theoretically necessary in this replication study, we will focus on the first step, and include fewer background variables for the second step. 

### Methods Addendum (Post Data Collection)

#### Actual Sample

The final sample after data exclusion only includes 150 people. Despite that I used the "full-time employee only" filter on Prolific, there are 20 people who do not fulfill this requirement in my further screening in the questionnaire. For example, over 10 people are self-employed but not employed by a company, some people are part-time employees or unemployed. In the future I should recruit a larger sample to take into account the lack of specificity in Prolific screening. 

Participants (aged 22-74, with a mean of 44.24) include 81 males, 68 females, and 1 non-binary individual. Their subjective socioeconomic status range from 1 to 9 (1 being the lowest, 10 being the highest), with a mean of 5.513. 

#### Differences from pre-data collection methods plan 

Besides the smaller sample size, everything else was consistent with the original plan.

## Results

### Data preparation
	
```{r Data preparation}

###Data Preparation

####Load Relevant Libraries and Functions

library(tidyverse)
library(knitr)
library(MASS)
library(nnet)
library(haven)
library(car)
library(psych)
library(RColorBrewer)
library(ggsignif)
library(DescTools)
library(rcompanion)
library(dplyr)

#### Import data

# Load and clean data
#setwd("~/Downloads/Psych251/FinalProject/shavit2022/writeup")
psych251data <- read.csv("psych251data_choicetext.csv", header = TRUE)
#View(psych251data)
names(psych251data)

#### Data exclusion / filtering

psych251data <- psych251data[-c(1:25),] # Remove first two rows meta-data + pilot participants
nrow(psych251data) # Confirm having 170 participants in the study
psych251data <- psych251data %>% dplyr::select(ResponseId, Unspecified, Expanded, Age, Gender, Employment, SESLadder) # Select relevant variables (deleting metadata)

#YOCHAI'S ADDITION: RECODE CHOICE FROM NUMBER TO TEXT
table(psych251data$Unspecified)#- Note: Majority Help colleague/ friend

psych251data$Unspecified=recode(psych251data$Unspecified, "
        'Help a colleague or friend with a company project'=1; 
        'Work on a company project that may help you advance your career'=3;
        'Work on a company project that may lead the company in a new direction'=2
", as.factor=F, as.numeric=T)

table(psych251data$Unspecified)

table(psych251data$Expanded)#-> Note: MAjority Advance career
psych251data$Expanded=recode(psych251data$Expanded, "
        'Help a colleague or friend with a company project'=1; 
        'Work on a company project that may help you advance your career'=3;
        'Work on a company project that may lead the company in a new direction'=2
", as.factor=F, as.numeric=T)

table(psych251data$Expanded)

# Make sure all data are in an appropriate format
psych251data$Unspecified <- as.numeric(psych251data$Unspecified)
psych251data$Expanded <- as.numeric(psych251data$Expanded)
psych251data$Age <- as.numeric(psych251data$Age)
psych251data$Gender <- factor(psych251data$Gender)
psych251data$Employment <- factor(psych251data$Employment)
psych251data$SESLadder <- as.numeric(psych251data$SESLadder)

table(psych251data$Employment)

psych251data <- psych251data %>% filter(Employment=="Working - as a full-time paid employee") %>% filter(Age >= 20 & Age <= 75) # We only want people aged 20-75 who are full-time paid employees.
summary(psych251data$Age)
summary(psych251data$Employment)
summary(psych251data$Gender)
summary(psych251data$SESLadder)

table(psych251data$Unspecified)#-> 41 help, 9 new direction, 27 advance career
table(psych251data$Expanded)#-> 17 help, 15 new direction, 41 advance career

#### Prepare data for analysis - create columns etc.

# Take a look at the data
psych251data <- as.data.frame(psych251data)

# Create a condition column to specify which condition each person is assigned to 
psych251data <- psych251data %>%
  mutate(Condition = case_when(
    Expanded == 1 | Expanded == 2 | Expanded == 3 ~ "Expanded",
    Unspecified == 1 | Unspecified == 2 | Unspecified == 3 ~ "Unspecified"))

psych251data=psych251data%>%group_by(ResponseId)%>%mutate(Condition2= 
      ifelse(is.na(Expanded)==F & is.na(Unspecified)==T,'Expanded',
             ifelse(is.na(Unspecified)==F & is.na(Expanded)==T, 'Unspecified', NA)))%>%ungroup()

psych251data$Condition==psych251data$Condition2 #-> Seems to be no difference between the two ways of creating a 'condition' variable

# Combine participants' answers to time horizon questions into one column called Choice
#psych251data <- gather(psych251data, "Condition", "Choice", c("Unspecified", "Expanded")) %>% filter(!is.na(Choice))

psych251data=psych251data%>%group_by(ResponseId)%>%mutate(Choice=ifelse(Condition=='Expanded', Expanded, ifelse(Condition=='Unspecified', Unspecified, NA)))%>%ungroup()

table(psych251data$Choice, psych251data$Condition)#-> 41 '1' (i.e. 'Help') in Unspecified & 17 '1' in Expanded; 27 '3' (i.e. 'Career') in Unspecified & 41 '3' in Expanded; 9 '2' (i.e 'New direction') in Unspecified & 15 '2' in Expanded. In all: 41 Help and 36 Future oriented in Unspecified; 17 Help and 56 Future oriented in Expanded

# Make sure choice and condition are factors 
psych251data$Choice_factor <- as.factor(psych251data$Choice)
psych251data$Choice_factor <- dplyr::recode(psych251data$Choice, '1' = "colleague/friend", '2' = "new direction", '3' = "advance career")

table(psych251data$Choice_factor, psych251data$Condition)#-> 41 'Colleague/friend' in Unspecified & 17 in Expanded; 27 'Advance career' in Unspecified & 41 in Expanded; 9 'New direction' in Unspecified 15 in Expanded. In all: 41 Help and 36 Future oriented in Unspecified; 17 Help and 56 Future oriented in Expanded

psych251data$Condition=as.factor(psych251data$Condition)
#levels(psych251data$Condition)=c("Unspecified", "Expanded")

table(psych251data$Choice_factor, psych251data$Condition) #-> !!!!!!!!!!!! NOTE (if line 204 is not commented out: NOW THE CONDITION LABELS FLIPPED AND THEY ARE EXACTLY THE OPPOSITE THAN THE VARIABLE CONDITION2 !!!!!!!!!!!!!!! I THINK IT'S DUE TO THE USE OF THE 'LEVELS()' FUNCTION. What I think happened- The value that comes first alphabetically ('Expanded') gets the level that comes first in the function ('Unspecified')

#Compare with
table(psych251data$Choice_factor, psych251data$Condition2)

# Is there sig diff between the number of people in each condition
conditiontable=table(psych251data$Condition)
conditiontable
t.test(conditiontable)#-> Not sure why t.test here. I think it should be chi-square contingency test
chisq.test(conditiontable)#-> n.s

# Age, gender, SES diff across conditions
summary(lm(data = psych251data, Age~Condition))#-> Not significant
summary(table(psych251data$Gender,psych251data$Condition))#-> Not significant
summary(lm(data = psych251data, SESLadder~Condition))#-> Not significant
```

### Confirmatory analysis

The analyses as specified in the analysis plan.  

```{r Confirmatory analysis}

# Create binary variable (that indicates whether the participants choose to help others or not) for logistic regression
psych251data$Help <- recode (psych251data$Choice_factor, "'colleague/friend'=1; 'advance career'=0; 'new direction'=0; else=NA", as.factor = F, as.numeric = T)

table(psych251data$Help, psych251data$Condition)#-> 41 'Help' and 36 'Future' in Unspecified, 17 'Help' and 56 'Future' in Expanded

# Main analyses: logistic regression - difference between control and expanded time horizon conditions

# Set offset to reflect that chance success is 33%
p_help=0.333
log_odds_help = log(p_help/(1-p_help))
log_odds_help

# Difference between conditions (age by condition) with no control variables
psych251data$Condition=as.factor(psych251data$Condition)
psych251data$Condition <- relevel(psych251data$Condition, ref="Unspecified")
glm_diff_unspec <- glm(data = psych251data, Help~scale(Age)*Condition, family = "binomial", offset = rep(log_odds_help, length(psych251data$ResponseId)))
summary(glm_diff_unspec) #-> Note that now there is a positive simple effect of age on the likelihood to choose 'Help' in the unspecified condition (beta= 0.4375, p=0.05) and a strong main effect of condition suggesting the the likelihood of choosing 'Help' (b=-1.36, p<.001) in the expanded condition is lower. The interaction is not significant, but in the expected direction (i.e suggesting that the positive effect of age is reduced in the Expanded condition)

psych251data$Condition <- relevel(psych251data$Condition, ref="Expanded")
glm_diff_expand <- glm(data = psych251data, Help~scale(Age)*Condition, family = "binomial", offset = rep(log_odds_help, length(psych251data$ResponseId)))
summary(glm_diff_expand) #-> No effect of age in the Expanded condition (beta= 0.17, p=0.57)

# Difference between conditions (age by condition) with control variable: SES
psych251data$Condition <- relevel(psych251data$Condition, ref="Unspecified")
glm_diff_unspec_ctl <- glm(data = psych251data, Help~scale(Age)*Condition + scale(SESLadder), family = "binomial", offset = rep(log_odds_help, length(psych251data$ResponseId)))
summary(glm_diff_unspec_ctl) #-> findings unaffected by the inclusion of gender

psych251data$Condition <- relevel(psych251data$Condition, ref="Expanded")
glm_diff_expand_ctl <- glm(data = psych251data, Help~scale(Age)*Condition + scale(SESLadder), family = "binomial", offset = rep(log_odds_help, length(psych251data$ResponseId)))
summary(glm_diff_expand_ctl) 

```

*Side-by-side graph with original graph is ideal here*

For original graph, see Figure 1 in the original paper: https://github.com/psych251/shavit2022/blob/main/original_paper/original_paper.pdf

Note that in this replication study, we only have unspecified time horizon condition and expanded time horizon condition. We did not include the limited time horizon condition.

```{r Graphs}

psych251data$Condition <- relevel(psych251data$Condition, ref="Unspecified")
psych251data$Condition_Order <- as.factor(psych251data$Condition)
levels(psych251data$Condition_Order) <- c("Unspecified", "Expanded time")
levels(psych251data$Condition_Order)
table(psych251data$Condition_Order)

ggplot(data=psych251data, aes(x=Age, y=Help)) +
  geom_point(alpha=0.2, position = position_jitter(height = 0.03)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), color="black") +
  geom_hline(yintercept = 0.33, linetype="dashed") +
  ggthemes::theme_few()+labs(y="Likelihood of choosing\n'Help friend / colleague'", x="Age") +
  facet_grid(.~Condition_Order)

```

### Exploratory analyses

```{r Exploratory analyses}

# Exploratory analyses: logistic regression - within control or within expanded time horizon conditions

# Within unspecified condition with no control variables

psych251data_unspec <- psych251data %>% filter(Condition=="Unspecified")
glm_unspec <- glm(data = psych251data_unspec, Help~scale(Age), family = "binomial", offset = rep(log_odds_help, length(psych251data_unspec$ResponseId)))
summary(glm_unspec)

# Within expanded condition with no control variables

psych251data_expand <- psych251data %>% filter(Condition=="Expanded")
glm_expand <- glm(data = psych251data_expand, Help~scale(Age), family = "binomial", offset = rep(log_odds_help, length(psych251data_expand$ResponseId)))
summary(glm_expand)

# Within unspecified condition with control variable: SES

psych251data_unspec <- psych251data %>% filter(Condition=="Unspecified")
glm_unspec_ctl <- glm(data = psych251data_unspec, Help~scale(Age) + scale(SESLadder), family = "binomial", offset = rep(log_odds_help, length(psych251data_unspec$ResponseId)))
summary(glm_unspec_ctl)

# Within expanded condition with control variable: SES

psych251data_expand <- psych251data %>% filter(Condition=="Expanded")
glm_expand_ctl <- glm(data = psych251data_expand, Help~scale(Age) + scale(SESLadder), family = "binomial", offset = rep(log_odds_help, length(psych251data_expand$ResponseId)))
summary(glm_expand_ctl)

```

## Discussion 

### Summary of Replication Attempt 

We replicated the original result. 

Exploratory analyses:

Like the original study, we found that older adults (compared to younger adults) are more likely to help others in the unspecified time horizon condition. Additionally, when we specify the time horizon to be expanded, age difference in choosing to help others will be eliminated. 

Main analyses: 

There is a significant difference between the unspecified and expanded time horizon condition in the age differences in work motivation, such that in the unspecified time horizon condition, older adults would be more likely to choose to help others than younger adults, but in the expanded time horizon condition, both older and younger adults are less likely to choose to help others, and there is no significant age difference. 

### Commentary

There are some differences between the replication study and the original study. For example, our replication study has only a sample size of 150, and therefore our analyses were under powered, while the original study has a sample of 555. Additionally, our replication could not include as many covariates as the original study did. However, for both my main analyses and exploratory analyses, the original study had significant findings both with and without covariates. Moreover, our replication study made up for a limitation of the original study, which included both full-time and part-time employees. The author of the original study mentioned that part-time employees may lean towards future-oriented goals because they may be more interested in pursuing a full-time job elsewhere, so in the replication study, we only included full-time employees. However, I don't believe that any of these differences between the replication study and the original study would significantly influence the results.

Several limitations of both the original and replication results include: 1. People may have different perceptions of these goals. For example, even though the original study used "helping others" to represent emotionally meaningful goals, to some new employees in a company, helping others may be associated with exploratory goals as it may help them build professional network and thereby facilitate their professional development. 2. Our experimental design may not be completely realistic for the participants. For example, in real life, many employees may not really have the choice of how to spend their time each day at work. Also, there are a lot of other work-related activities while our study design restricts their choices to three activities. These are some possible reasons why the results did not replicate.  

A challenge that I will keep in mind in the future is sample size estimation. Even though Prolific has its own screener, when I used my own screener in Qualtrics survey, I screened out 20 participants (out of 170). In the future I should recruit more participants than the number that my power analysis suggests to buffer any participant loss during screening.
